---
title: 'Solution for Dataquest Guided Project: Exploring NYC Schools Survey Data'
author: "Huseyn Rahimov"
date: "April 21 2019"
output: html_document 
---

**Here is the suggested solution in the Data Cleaning With R Guided Project: Exploring NYC Schools Survey Data.**

Loading the packages we will need for our analysis

```{r}
library(readr)
library(dplyr)
library(stringr)
library(purrr)
library(tidyr)
library(ggplot2)
```

Import the data into R (rsv won't work for "survey and survey_d75"" data here, use read_tsv)
```{r}
combined <- read_csv("combined.csv") 
survey <- read_tsv("masterfile11_gened_final.txt")
survey_d75 <- read_tsv("masterfile11_d75_final.txt")
```

Filtering `survey` data to include only high schools and selecting columns that are imprtant for our analysis
```{r}
survey_select <- survey %>% filter(schooltype == "High School") %>% select(dbn:aca_tot_11)
```

Select columns needed for analysis from `survey_d75`, we dont need to select schools as all are the same
```{r}
survey_d75_select <- survey_d75 %>% select(dbn:aca_tot_11)
```

let's combine `survey` and `survey_d75` data frames
```{r}
survey_total <- survey_select %>% bind_rows(survey_d75_select)
```

let's rename the `dbn` to `DBN` in `survey_total` data frame, so that we can join with `combined`
```{r}
survey_total <- survey_total %>% rename(DBN = dbn)
```

let's combine `survey_total` and `combined` data frames now by using left_join, it keeps the survey that that is related with `combined` data frame
```{r}
combined_survey <- combined %>% left_join(survey_total, by = "DBN")
```

We can anytime view the first few rows of the any data frame after making changes, that is good to see the track of changes (what we have done) or just from the Global Environment on the top right-hand side click the created Data names and it will show up in the R script(better view)
```{r}
head(combined_survey)
```

Now let's look at the correlation between variables, we will choose the ones that we feel necessary for our analysis and coverting to "Tibble" which we will need later in scatter plots for selecting variables
```{r}
cor_mat <- combined_survey %>% select(avg_sat_score, saf_p_11:aca_tot_11) %>% cor(use = "pairwise.complete.obs")

cor_tib <- cor_mat %>% as_tibble(rownames = "variable")
```

Looking for correlations with `avg_sat_score`, < -0.25 and > 0.25 which are strong ones
```{r}
strong_cors <- cor_tib %>% select(variable, avg_sat_score) %>% filter(avg_sat_score > 0.25 | avg_sat_score < -0.25)
```

Scatter plots of these variables with `avg_sat_score`
```{r}
create_scatter <- function(x, y) {ggplot(data = combined_survey) + aes_string(x = x, y = y) + geom_point(alpha = 0.3) + theme(panel.background = element_rect(fill = "white"))}

x_var <- strong_cors$variable[2:5]
y_var <- "avg_sat_score"
map2(x_var, y_var, create_scatter)
```

Reshaping the data to see differences in "student", "parent" and "teacher" responses and so creating 2 new columns - "survey_question" and "score"
```{r}
combined_survey_gather <- combined_survey %>% gather(key = "survey_question", value = score, saf_p_11:aca_tot_11)
```

from Survey_question variables we also need to extract above mentioned responses and also the type of question
```{r}
combined_survey_gather <- combined_survey_gather %>% mutate(response_type = str_sub(survey_question, 4, 6)) %>% mutate(question = str_sub(survey_question, 1, 3))
```

renaming the Response type column with its respective names
```{r}
combined_survey_gather <- combined_survey_gather %>% mutate(response_type = ifelse(response_type == "_p_", "parent", ifelse(response_type == "_t_", "teacher", ifelse(response_type == "_s_", "student", ifelse(response_type == "_to", "total", "NA")))))
```

Creating a box plot(excluding Total), looking at relations between the responses, question types and relative scores
```{r}
combined_survey_gather %>% filter(response_type != "total") %>% ggplot() + aes(x = question, y = score, fill = response_type) + geom_boxplot()
```

